---
title: "P8105 HW6"
author: Chenxin Zhang
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%", 
  fig.align = 'center',
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  echo = TRUE
)
theme_set(theme_minimal() + 
          theme(legend.position = "bottom") +
          theme(plot.title = element_text(hjust = 0.5, size = 15, face = "bold")) +
          theme(plot.subtitle = element_text(hjust = 0.5))
          )
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Problem 1

```{r load_tidy_data1}
homicide_df =
  read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

Start with one city.

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")
# fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate), # odds ratio
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

Try this across cities.

```{r}
models_results_df =
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
    ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate), # odds ratio
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Problem 2
**Import and Tidy Data**
  *`babysex:` baby’s sex (male = 1, female = 2)
  *`frace:` father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
  *`mrace:` mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
  *`malform:` presence of malformations that could affect weight (0 = absent, 1 = present)
  
```{r load_tidy_data2}
#load draft data frame
#convert numeric to factor
#change pounds to grams
bw_df = read_csv("./data/birthweight.csv") %>% janitor::clean_names() %>% 
  mutate(
    babysex = factor(babysex, labels = c("male", "female")),
    frace = factor(frace, levels = c(1,2,3,4,8,9), labels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
    mrace = factor(mrace, levels = c(1,2,3,4,8,9), labels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
    malform = factor(malform, labels = c("absent", "present")),
  delwt = delwt * 453.59237,
    ppwt = ppwt * 453.59237,
    wtgain = wtgain * 453.59237) %>% 
  arrange(babysex, frace)
```

```{r check_missing_data}
#there are two ways to check for missing data
sum(is.na(bw_df))
which(is.na(bw_df))
```

* The dataset is consist of `r nrow(bw_df)` observations and `r ncol(bw_df)` variables, and there is `r sum(is.na(bw_df))` missing value. The numeric variable `babysex`, `frace`, `mrace`, and `malform` are converted to factor variable.

**Propose a regression model for birthweight**

I am interested in how factors associated with mother can affect baby's birthweight, so I will choose   
`delwt:`(mother’s weight at delivery,   
`mheigth:` mother’s height (inches),   
`momage:` mother’s age at delivery (years),
`smoken:` average number of cigarettes smoked per day during pregnancy as predictors to build linear regression model. 

```{r}
my_model = bw_df %>% 
  lm(bwt ~ delwt + mheight + momage + smoken, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable()
my_model
```
* The results show a positive intercept of 978.977072, and the parameters of delwt, mheight, momage and smoke are 0.012720, 15.969731, 15.304540, -7.023537 respectively. The effect of these four predictors on birthweight are significant, since the P-value of all predictors are much less than 0.05.

**model residuals against fitted values**

```{r}
my_model = bw_df %>% 
  lm(bwt ~ delwt + mheight + momage + smoken, data = .) 
plot = bw_df %>% 
  add_predictions(my_model) %>% 
  add_residuals(my_model) %>% 
  rename("prediction" = pred, "residual" = resid) %>%
  ggplot(aes(x = prediction, y = residual)) +
    geom_point(size = 0.9, alpha = 0.5, color = "pink") +
    labs(title = "Plot of Residual vs prediction", x="Predicted/Fitted value(grams)", y="Residual(grames)") + 
    geom_abline(intercept = 0, slope = 0)
plot
```
* we can not see obvious relationship between the residuals and the predicted values from the plot above.

**Compare the three models**

* model1 using length at birth and gestational age as predictors (main effects only)

```{r}
model1 = lm(bwt ~ blength + gaweeks, data = bw_df) %>% broom::tidy() %>% 
  knitr::kable()
model1
```

* model2 using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
model2 = lm(bwt ~ bhead * blength * babysex, data = bw_df) %>% broom::tidy() %>% 
  knitr::kable()
model2
```
**cross validation**

```{r}
#crossv_mc() will get 500 repeated sampling, each sample has sample size as the original dataset, and create train and test data
cv_df = 
  crossv_mc(bw_df, 500) %>% 
    mutate(
        train = map(train, as.tibble),
        test = map(test,as.tibble)
    )  %>%
  
  mutate(my_model  = map(train, ~lm(bwt ~ delwt + mheight + momage + smoken, data = .x)),
         model1  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         model2  = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))) %>% 
  
  mutate(rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
         rmse_model1 = map2_dbl(model1 , test, ~rmse(model = .x, data = .y)),
         rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x, data = .y))) 
cv_df
```


```{r}
#select interested variables 
rmse_df = cv_df %>% 
select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>%   mutate(
    model = fct_reorder(model,rmse))
rmse_df
```

```{r}
plot2 = rmse_df %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin(aes(fill = model), color = "black", alpha = .5) +
    labs(title = "Comparisions of three models", y="RMSE value", x = "Models") 
plot2
```
* From the plot, we can observe that `my_model` has greater RMSE than the other two models, and `model2` has smaller RMSE than `model1`. The results peove that prediction accuracy of `model2` is higher than the other two. The prediction accuracy of `my_model` is too low to be a good linear model. ter than `model1`. I would recommand use `model2` for  hypothetical analysis, or find a better model with higher prediction accuracy.
